---
title: "Project2"
author: "utkrsh"
date: "2024-05-13"
output:
  pdf_document: default
---
# Importing required libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)    # ggplot(), %>%, mutate(), and friends
library(scales)       # Format numbers with functions like comma(), percent(), and dollar()
library(broom)        # Convert models to data frames
library(modelsummary) # Side-by-side regression tables
library(foreign)   # for importing stata data
library(readr)
library(haven)
library(modelsummary)
library(stargazer)
library(tidyverse)
```
# Importing dataset "INJURY" 
```{r}
dataset <- read_dta("C:/Users/utkrsh/Desktop/EC402_Assignment_2_DID/INJURY.DTA")
```
# Cleaning the data so that it only includes rows from (ky == 1)

```{r}
injury <- dataset %>% filter(ky==1)


```

# Renaming the columns

```{r}
injury <- injury %>% rename(duration = durat, log_duration = ldurat,after_1980 = afchnge)
```
Viewing the dataset
```{r}
injury
```

# Converting "industry" and "injury type" to categories/factors
```{r}
df <- injury %>% mutate(indust = as.factor(indust),injtype = as.factor(injtype))

```

## 1.	Calculating the policy effect on duration, without running any regression, here we see the mean duration of the weeks for both treated and control group before and after 1980 (treatement)
```{r}
difr <- df %>% group_by(after_1980,highearn) %>% summarize(mean.duration = mean(duration))
print(difr)
```
### [after_1980(0);highearn(0)]: pre-treatement control group  mean_duration    : 6.47  weeks
### [after_1980(0);highearn(1)]: pre-treatement treatement group mean_duration  : 11.76 weeks
### [after_1980(1);highearn(0)]: post-treatement control group mean_duration    : 7.03  weeks
### [after_1980(1);highearn(1)]: post-treatement treatement group mean_duration : 12.89 weeks


### policy_effect =[(avg_duration(post-treatement treated group)-avg_duration(pre-treatement treated group)]-[avg_duration(post-treatement control group)-avg_duration(pre-treatement control group)]

### policy_effect = [avg_duration of treated group (POST-PRE)]-[avg_duration of control group(POST-PRE)]
```{r}
pre_treatement_treated_group <- difr %>% 
  filter(after_1980 == 0, highearn == 1) %>% 
  pull(mean.duration)

pre_treatement_control_group<- difr %>% 
  filter(after_1980 == 0, highearn == 0) %>% 
  pull(mean.duration)

post_treatement_treated_group <- difr %>% 
  filter(after_1980 == 1, highearn == 1) %>% 
  pull(mean.duration)

post_treatement_control_group <- difr %>% 
  filter(after_1980 == 1, highearn == 0) %>% 
  pull(mean.duration)
```

```{r}
treatement_group_before_after <- post_treatement_treated_group - pre_treatement_treated_group
control_group_before_after    <-post_treatement_control_group- pre_treatement_control_group
```




## Policy Effect(DiD Estimate:)
```{r}
policy_effect_nive <- treatement_group_before_after - control_group_before_after
print(policy_effect_nive)
```


## 2.	Calculating the policy effect on duration, without running any regression, here we see the mean log_duration in  weeks for both treated and control group before and after 1980(treatemnt)

```{r}
difr_log <- df %>% group_by(after_1980,highearn) %>% summarize(mean.log_duration = mean(log_duration))
head(difr_log)
```

### [after_1980(0);highearn(0)]: pre-treatement control group  mean_log_duration    : 1.12 weeks
### [after_1980(0);highearn(1)]: pre-treatement treated group mean_log_duration     : 1.38 weeks
### [after_1980(1);highearn(0)]: post-treatement control group mean_log_duration    : 1.13 weeks
### [after_1980(1);highearn(1)]: post-treatement treated group mean_log_duration    : 1.58 weeks

### policy_effect =[(log_avg_duration(post-treatement treated group)-log_avg_duration(pre-treatement treated group)]-[log_avg_duration(post-treatement control group)-log_avg_duration(pre-treatement control group)] ###

### policy_effect = [log_avg_duration of treated group (POST-PRE)]-[log_avg_duration of control group(POST-PRE)]

```{r}
pre_treatement_treated_group_log <- difr_log %>% 
  filter(after_1980 == 0, highearn == 1) %>% 
  pull(mean.log_duration)

pre_treatement_control_group_log<- difr_log %>% 
  filter(after_1980 == 0, highearn == 0) %>% 
  pull(mean.log_duration)

post_treatement_treated_group_log <- difr_log %>% 
  filter(after_1980 == 1, highearn == 1) %>% 
  pull(mean.log_duration)

post_treatement_control_group_log <- difr_log %>% 
  filter(after_1980 == 1, highearn == 0) %>% 
  pull(mean.log_duration)
```

```{r}
log_treatement_group_before_after <- post_treatement_treated_group_log - 
  pre_treatement_treated_group_log
log_control_group_before_after    <-post_treatement_control_group_log- 
  pre_treatement_control_group_log
```




## Policy Effect log(DiD log Estimate:)
```{r}
policy_effect_log <- log_treatement_group_before_after - log_control_group_before_after
print(policy_effect_log)
```


# Plotting 
### Distribution of Duration by Category
```{r}
ggplot(data = df, aes(x = duration)) +
  geom_histogram(binwidth = 10, color = "white", boundary = 0) +
  facet_wrap(~ highearn)
```

### Ploting log duration based on category
```{r}
ggplot(data = df, mapping = aes(x = log_duration)) +
  geom_histogram(binwidth = 0.5, color = "white", boundary = 0) + 
   scale_x_continuous(labels = trans_format("exp", format = round)) +
  facet_wrap(~ highearn)
```

### Here we just calculate Mean
```{r}
ggplot(df, aes(x = factor(highearn), y = log_duration)) +
  geom_point(size = 0.5, alpha = 0.2) +
  stat_summary(geom = "point", fun = "mean", size = 5, color = "blue") +
  facet_wrap(vars(after_1980))
```









## 3.Basic Regression analysis to calculate the estimates without any control 
### Basic Regression model
$$
duration = \beta_0 + \beta_1 after1980 + \beta_2 highearn + \delta afterchange \cdot highearn + u
$$


```{r}
basic_model_1 <- lm(duration ~ after_1980 + highearn + after_1980*highearn, data = df)
tidy(basic_model_1)
```
The notation $\delta$$1$ = 0.9513, indicates that the policy change might have increased the duration of benefits for
high-income workers by about 0.9513 weeks more than for low-income workers. However, this effect is not
statistically significant (p=0.414). The coefficient on after_1980 is small 0.7658 and statistically insignificant
which means the increase in the earnings cap has no effect on duration for low-income workers.


$$
\log(duration) = \beta_0 + \beta_1 after1980 + \beta_2 highearn + \delta_1 afterchange \cdot highearn + u
$$

```{r}
basic_model_2 <- lm(log(duration) ~ after_1980 + highearn + after_1980*highearn, data = df)
tidy(basic_model_2)
```

The notation $\delta$ signifies that the average duration of workers' compensation among high earners rose approximately by 19.06% due to the increased earnings cap. The coefficient on after_1980 is small 0.007 and statistically insignificant which means the increase in the earnings cap has no effect on duration for low-income workers.

## 4.regression adujstment procedure to evaluate the impact of policy change on the duration, and log_duration.

### Regression Adjustment Model 1
$$
\text{duration} = \beta_0 + \beta_1 \text{after1980} + \beta_2 \text{highearn} + \delta_1 \text{afterchange} \cdot \text{highearn} \\$$
$$+ \gamma_1\text{male} + \gamma_2\text{married} + \gamma_3\text{age}^2 + \gamma_4\text{hosp} + \gamma_5\text{indust} \\$$
$$+ \gamma_6\text{injtype} + \gamma_7\text{lprewage}+ u$$
### Creating a new column age squared
```{r}
df <- df %>% mutate(age_squared = age^2)

```

```{r}
adv_model_1 <- lm(duration ~ after_1980 + highearn + 
after_1980*highearn + male + married + age_squared + 
hosp + indust + injtype + lprewage, data =df)
print(summary(adv_model_1))
```
After controlling for all the X's (male ,married , age_squared ,hosp,indust,injtype,lprewage), the coeffcicent of interation term comes out to be 1.79 whic is insignificant. 

### Regression Adjustment Model 2

$$
\log(\text{duration}) = \beta_0 + \beta_1 \text{after1980} + \beta_2 \text{highearn} + \delta_1 \text{afterchange} \cdot \text{highearn} \\$$
$$
+ \gamma_1\text{male} + \gamma_2\text{married} + \gamma_3\text{age}^2 + \gamma_4\text{hosp} + \gamma_5\text{indust} \\$$
$$+ \gamma_6\text{injtype} + \gamma_7\text{lprewage}+ u$$

After controlling for all the X's (male ,married , age_squared ,hosp,indust,
injtype,lprewage), the coeffcicent of interation term comes out to be 1.69 whic 
is significant , which means that the he average duration of workers' 
compensation among high earners rose approximately by 16.07% due to the 
increased earnings cap. 




```{r message=FALSE, warning=FALSE}
adv_model_2 <- lm(log_duration ~ after_1980 + highearn + 
after_1980*highearn + male + married + age_squared + hosp 
+ indust + injtype + lprewage, data =df)
print(summary(adv_model_2))
```




Combining the data set to contain (ky==0)
We here combine the data set including the state of michigan state sample . 
```{r}
dataset_combined <- dataset %>% rename(duration = durat, 
log_duration = ldurat,after_1980 = afchnge)%>% 
mutate(indust = as.factor(indust),injtype = as.factor(injtype))%>% 
mutate(age_squared = age^2)
```

# 5.  Robustness check
$$\log(\text{duration}) = \beta_0 + \beta_1 \text{after1980} + \beta_2 \text{highearn}
+ \delta_1 \text{afterchange} \cdot \text{highearn} \\$$
$$+ \gamma_1\text{male} + \gamma_2\text{married} + \gamma_3\text{age}^2 \\
+ \gamma_4\text{hosp} + \gamma_5\text{indust} \\$$
$$+ \gamma_6\text{injtype} + \gamma_7\text{lprewage} \\
+ \delta_2 \text{after1980} \cdot \text{ky} + \delta_3 \text{ky} + u
$$
```{r}
robust_model1 <- lm(log_duration ~ after_1980 + highearn +
after_1980*highearn + male + married + age_squared + hosp 
+ indust + injtype + lprewage + ky +ky:after_1980, data =dataset_combined)
print(summary(robust_model1))
```
## DDD

$$
\text{log(duration)}= \beta_0+ \beta_1*ky+ \beta_2*after1980_t+ \beta_3*highEarn_i+\\
$$
$$\gamma_1(ky_s*after1980_t)+\gamma_2(ky*highearn)+ \gamma_2(after1980*highearn)+\\
\delta_1(ky*after1980*highearn)+ \psi*{X}+ u$$


Considering the entire dataset including Michigan, we need to account for 
state-specific and income-specific time trends that might influence compensation
benefits. To do this, we use Triple Differences, which involves comparing the 
differences between Kentucky and Michigan data to eliminate these confounding 
factors.
```{r}
DDD= lm( duration ~ ky + after_1980 + highearn + ky*after_1980 + ky*highearn 
         + after_1980*highearn + ky*after_1980*highearn + male + married 
         + age_squared + hosp 
+ indust + injtype + lprewage, data=dataset_combined)
summary(DDD)

```
here we see that the triple interaction term comes out to be 1.36 positive but 
insignificant, when controlling for all the other possible X's.



# 6. Generating Graph
```{r}
graph_data <- dataset_combined %>% 
   mutate(highearn = factor(highearn, labels = c("Low earner", "High earner")),
         after_1980 = factor(after_1980, labels = c("Before 1980", "After 1980"))) %>% 
  group_by(highearn, after_1980) %>% 
  summarize(mean_duration = mean(log_duration),
            se_duration = sd(log_duration) / sqrt(n()),
            upper = mean_duration + (1.96 * se_duration),
            lower = mean_duration + (-1.96 * se_duration)) 
```
```{r}
ggplot(graph_data, aes(x = highearn, y = mean_duration)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                  color = "darkorange", size = 1) +
  facet_wrap(vars(after_1980))
```

```{r}
ggplot(graph_data, aes(x = after_1980, y = mean_duration, color = highearn)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), size = 1) + 
  # The group = highearn here makes it so the lines go across categories
  geom_line(aes(group = highearn))
```

```{r}
ggplot(difr_log, aes(x = as.factor(after_1980), 
                  y = mean.log_duration, 
                  color = as.factor(highearn)))+
  geom_point() +
  geom_line(aes(group = as.factor(highearn))) +
  annotate(geom = "segment", x = "0", xend = "1",
           y = pre_treatement_treated_group_log, yend = 
             post_treatement_treated_group_log - policy_effect_log,
           linetype = "dashed", color = "black") +
  annotate(geom = "segment", x = "1", xend = "1",
           y = post_treatement_treated_group_log, yend =
             post_treatement_treated_group_log - policy_effect_log,
           linetype = "dotted", color = "red") +
  annotate(geom = "label", x = "1", y = post_treatement_treated_group_log
           - (policy_effect_log / 2), 
           label = "Program effect", size = 5)
```

```{r message=FALSE, warning=FALSE}
modelsummary(list(basic_model_2,adv_model_2,DDD,robust_model1),output='latex')

```
